Computing volume via non-linear rectangular prisms

### 2019-07-30 Cav2D/article-template.tex
added starting math for cavalieri parametric 2d

@@

>>>>

%MAKING THE CAVALIERI INTEGRAL NOTATION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\makeatletter
%\let\DOTSI\relax % amsmath support for \dots
\newcommand*{\cint}{%
  \DOTSI
  \mathop{%
    \mathpalette\@LetterOnInt{\mathbf{c}}%
  }%
  \mkern-\thinmuskip % thin space is inserted between two \mathop
  \int
}
\newcommand*{\@LetterOnInt}[2]{%
  \sbox0{$#1\int\m@th$}%
  \sbox2{$%
    \ifx#1\displaystyle
      \textstyle
    \else
      \scriptscriptstyle
    \fi
    #2%
  \m@th$}%
  \dimen@=.4\dimexpr\ht0+\dp0\relax
  \ifdim\dimexpr\ht2+\dp2\relax>\dimen@
    \sbox2{\resizebox*{!}{\dimen@}{\unhcopy2}}%
  \fi
  \dimen@=\wd0 %
  \ifdim\wd2>\dimen@
    \dimen@=\wd2 %
  \fi
  \rlap{\hbox to \dimen@{\hfil
    $#1\vcenter{\copy2}\m@th$%
  \hfil}}%
  \ifdim\dimen@>\wd0 %
    \kern.5\dimexpr\dimen@-\wd0\relax
  \fi
}
\makeatother
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\section{Cavalieri Integral}
In this section we define the Cavalieri integral \cite{ackermann12}.

Let the region $R$ be bounded by the $x$-axis and the lines $f(x)=x$, $a(y)=1-y$ and $b(y)=4-y$. This region is shown in Fig.~\ref{fig:caval2}.\\
%\begin{figure}[htb]
%\centering
%\includegraphics[width=0.3\textwidth]{fig12}
%\caption{Region bounded by the $x$-axis and the lines $f(x)=x$, $a(y)=1-y$, and $b(y)=4-y$. Reproduced from Quaestiones Mathematicae (2012) 35: 265-296 with permission \copyright~ NISC (Pty) Ltd.}
%\label{fig:ex1}
%\end{figure}

\begin{figure}[htb]
\centering
\includegraphics[width=0.75\textwidth]{fig13.eps}
\caption{Region bounded by the $x$-axis and the lines $f(x)=x$, $a(y)=1-y$, and $b(y)=4-y$. The figure also depicts the partition points $x_i^1$ and $x_i^2$ as used in the Cavalieri sum (see Eq.~\eqref{eq:cav_sum}). Reproduced from Quaestiones Mathematicae (2012) 35: 265-296 with permission \copyright~ NISC (Pty) Ltd.}
\label{fig:caval2}
\end{figure}

\noindent
We can compute the area of $R$ as follows:
\begin{equation}
\int_0^2x\, dx+\int_2^44-x\, dx- \int_0^{\frac{1}{2}}x\, dx-\int_{\frac{1}{2}}^11-x\, dx = 3.75. 
\end{equation}

\noindent
Alternatively, the area of $R$ can also be determined via the Cavalieri integral:
\begin{equation}
\label{eq:caval1}
\int_{a(y)}^{b(y)}f(x)\, dx = \lim_{n\to \infty}\sum_{i=0}^{n-1} f(x_i^2)\Delta x_i^1,
\end{equation}
with $\Delta x_i^1 = x_{i+1}^1 - x_i^1$. Furthermore, the partition points $(x_i^1)_{i=0}^{n}$ and $(x_i^2)_{i=0}^{n}$ are depicted in Fig.~\ref{fig:caval2}. Let us review why this is in fact the case:
\begin{itemize}
 \item $\Delta x_i^1 = x_{i+1}^1 - x_i^1$ represents the width of a non-rectangular integration strip.
 \item $f(x_i^2)$ represents the height of a non-rectangular integration strip.
 \item $f(x_i^2)\Delta x_i^1$ represents the height of a non-rectangular integration strip (due to Cavalieri's principle).
 \item The Cavalieri sum $\sum_{i=0}^{n-1} f(x_i^2)\Delta x_i^1$, therefore, approximates the area of $R$.
 \item In the limit the Cavalieri sum, therefore, approaches the area of $R$.
\end{itemize}

\noindent
The Cavalieri integral is, however, hard to evaluate directly. Fortunately, it is easy to convert a Cavalieri integral into an equivalent Riemann integral:
\begin{equation}
\int_{a(y)}^{b(y)}f(x)\, dx = \int_{a}^{b} f\circ h(x)~dx, 
\end{equation}
with $a(0) = a$, $b(0)=b$ and $h = g^{-1}$. Furthermore, $g(x) = x - a\circ f(x) + a$.

\noindent
The area of $R$ can, therefore, also be computed as follows:
\begin{equation}
\int_{a(y)}^{b(y)}f(x)\, dx = \int_a^b f \circ h (x)\, dx = \dfrac{1}{2}\int_1^4x\, dx = 3.75.  
\end{equation}

\noindent
Expressing \eqref{eq:caval1} in vector calculus notation

\begin{equation}
\int_{a(y)}^{b(y)}f(x)\, dx =\cint_a^b f(x)~dx, 
\end{equation}
with $\mathbf{c}(t) = <c_x(t),c_y(t)>$ denoting a vector equation that passes through the origin and is a translation of $a(y)$. We can express $a(y)$ as $\mathbf{c}(t) + <a,0>$.
Note that partition $x_i^1$ and $x_i^2$ are related to each other via: 
\begin{equation}
\label{eq:pat_rel}
<x_i^2,f(x_i^2)> = <x_i^1,0> + <c_x(t),c_y(t)> 
\end{equation}

It follows from Eq.~\eqref{eq:pat_rel} that 
\begin{equation}
c_y(t) = f(x_i^1 + c_x(t)). 
\end{equation}
We can therefore express $t$ as a function of $x_i^1$. We can now compute $h(x_i^1) = x_i^1 + t(x_i^1)$.
